### 2024-04-02 paper/paper.qmd
Added placeholder abstract.

---
title: "Widening income gaps and the shrinking of the middle class"
subtitle: "An analysis of the impact of economic policies on the living standards of middle and lower-income households amidst rising costs and affordability challenges."
author: 
  - Abbass Sleiman
thanks: "Code and data are available at: [https://github.com/AbbassSleiman/US-Middle-Class](https://github.com/AbbassSleiman/US-Middle-Class)."
date: today
date-format: long
format: 
  pdf: 
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    documentclass: article
    geometry: margin = 1.2in
    abstract: "Over the past five decades, the United States has witnessed a decline in the percentage of adults residing in middle-class households, alongside escalating concerns about housing affordability. This study aims to assess the impact of rising costs on the living standards of middle and lower-income households, recognizing them as the most affected demographic by these economic shifts. Utilizing data from [...], we analyze changes in income distribution and evaluate the influence of economic policy changes, including interest rates and taxation, on the standards of living of these individuals. Additionally, we explore how individual-level characteristics intersect with these policies to address how best to pursue the goal of economic equity across individuals of all backgrounds. We find that [...]."
    number-sections: true
output: 
  bookdown::pdf_document2
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

The last 5 decades in the United States saw a clear contraction in the portion of adults residing in middle-class households - a steady decrease from 61% in 1971 to 50% in 2021 [@citepewmiddle]. Concerns regarding affordability of housing has increased drastically in recent years, as evidenced by the fact that 49% of Americans in 2021 found that the availability of affordable housing was a significant challenge [@citepewhouseaffordability]. Between 2017 and 2022, rents rose 18% and the average cost of all goods and services increased 16% [@citepewhouseaffordability].

Given that the middle class continues to comprise the majority of US adults, it is imperative to consider how the standards of living of these individuals, as well as those of lower-income households, have been impacted by these rising costs. As they are the most affected demographic by these economic shifts, understanding the implications on their daily lives, financial stability, and overall well-being is crucial for crafting effective policies and interventions to address these challenges.

In order to best understand the data and literature at hand, it is important to abide by a concrete definition of each of the three primary income brackets, lower-income, middle-income, and upper-income, to ensure consistency throughout the paper, as well as across the various literature cited. Following the definition provided by the Pew Research Center, this paper defines a middle-income adult in a given year as having an annual household income between two-thirds and twice the national median income in the year before, after adjusting for household size [@citepewmiddle]. Consequently, lower-income adults and upper-income adults are defined as those with less than two-thirds and over twice the median income respectively. It is important to note that we adjust for household size to account for the fact that smaller households require a lesser income to support similar lifestyles as larger households.

This paper utilizes data from [...] and seeks to examine how the relative proportions of each income class have changed over the past [...] years, and subsequently evaluate how various economic policy changes - interest rates, taxation, immigration, [...] - have influenced these proportions over time whilst offering a comprehensive analysis of how and why these policies may have caused the changes observed. Moreover, this paper aims to uncover how various individual-level characteristics including race, age, education, and [others], intersect with these policies to shape income distribution dynamics in order to address potential disparities and better explore how best to pursue the goal of economic equity across individuals of all
backgrounds.

This paper finds evidence that [...].

The remainder of this paper is structured as follows. [...].

# Data {#sec-data}

# Model

## Model set-up

### Model justification

# Results

# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

\newpage

# References